### 2018-02-09 article.tex
Put references in their own file

\documentclass[a4paper]{article}

\begin{document}

\end{document}