### 2018-02-09 article.tex
Added input metacontent

\documentclass[a4paper]{article}

\input{preamble}

\begin{document}

\input{metacontent}

\end{document}