### 2018-02-09 article.tex
Added section inputs

\documentclass[a4paper]{article}

\input{preamble}

\begin{document}

\input{metacontent}

\input{sections/1-introduction}
\input{sections/2-multiplicativefunctions}
\input{sections/3-adamsalgebras}
\input{sections/4-adamsalgebrasonmultfunc}
\input{sections/5-rationalmultfunc}
\input{sections/6-belltopology}
\input{sections/7-resultsonadamsalgebras}
\input{sections/8-identitiesapproach}
\input{sections/9-symmetricoperations}
\input{sections/10-algebraicgeometry}
\input{sections/11-dirichetletseries}
\input{sections/12-functorial}
\input{sections/13-lambdaringsclassicalfunctions}
\input{sections/14-classesmultfunc}
\input{sections/15-moreexamples}
\input{sections/16-WIP}
\end{document}