### 2018-02-12 article.tex
Renamed n-... to 0n-... so that lexical sorting is consistent

\documentclass[a4paper]{article}

\input{preamble}

\begin{document}

% title page, abstract, table of contents
\input{metacontent}

% chapters
\input{sections/01-introduction}
\input{sections/02-multiplicativefunctions}
\input{sections/03-adamsalgebras}
\input{sections/04-adamsalgebrasonmultfunc}
\input{sections/05-rationalmultfunc}
\input{sections/06-belltopology}
\input{sections/07-resultsonadamsalgebras}
\input{sections/08-identitiesapproach}
\input{sections/09-symmetricoperations}
\input{sections/10-algebraicgeometry}
\input{sections/11-dirichetletseries}
\input{sections/12-functorial}
\input{sections/13-lambdaringsclassicalfunctions}
\input{sections/14-classesmultfunc}
\input{sections/15-moreexamples}
\input{sections/16-WIP}

% references
\input{references}

\end{document}