### 2018-02-12 article.tex
Replaced inputs with include

\documentclass[a4paper]{article}

\input{preamble}

\begin{document}

% title page, abstract, table of contents
\include{metacontent}

% chapters
\include{sections/01-introduction}
\include{sections/02-multiplicativefunctions}
\include{sections/03-adamsalgebras}
\include{sections/04-adamsalgebrasonmultfunc}
\include{sections/05-rationalmultfunc}
\include{sections/06-belltopology}
\include{sections/07-resultsonadamsalgebras}
\include{sections/08-identitiesapproach}
\include{sections/09-symmetricoperations}
\include{sections/10-algebraicgeometry}
\include{sections/11-dirichetletseries}
\include{sections/12-functorial}
\include{sections/13-lambdaringsclassicalfunctions}
\include{sections/14-classesmultfunc}
\include{sections/15-moreexamples}
\include{sections/16-WIP}

% references
\include{references}

\end{document}