### 2018-02-16 sections/diagrams/perspective.tex

\newpage
\thispagestyle{empty}
\adjustbox{center}{
\begin{tikzpicture}[block/.style={draw=black,rounded corners=2ex,minimum width=7cm,minimum height=7cm,outer sep=5pt,inner sep=10pt},content/.style={inner sep=10pt,align=left,below right,text width=7cm},title/.style={align=center,below,font=\large,text height=0.5cm}]

  \node(atcenter) at (0,0) {}; 
  \node(multratbox)   [block,right=1cm of atcenter]{};
  \node(multratcircle)[block,left=1cm  of atcenter]{};
  \node(varieties)    [block,above=2cm of multratbox,minimum height=5cm]{};
  \node(lfunc)        [block,above=2cm of multratcircle,minimum height=5cm]{};
  \node(multcbox)     [block,below=2cm of multratbox,minimum height=6cm]{};
  \node(multccircle)  [block,below=2cm of multratcircle, minimum height=6cm]{};

  \node[title]   at (varieties.north){\uline{Varieties}};
  \node[content] at (varieties.north west){\ \\ \ \\ \ \\
  Varieties \\
  I'm right here!
  };

  \node[title]   at (lfunc.north){\uline{L-functions}};
  \node[content] at (lfunc.north west){\ \\ \ \\ \ \\
  L-functions \\
  I'm over here!
  };

  \node[title]   at (multratcircle.north){\uline{$Mult_{rat}(\C)$}};
  \node[content] at (multratcircle.north west){\ \\ \ \\ \ \\
  MultRat \\
  I'm just here!
  \begin{itemize}
  \item test1 \\
  \item test2
  \end{itemize}
  };

  \node[title]   at (multratbox.north){\uline{$Mult_{rat}(\C)$}};
  \node[content] at (multratbox.north west){\ \\ \ \\ \ \\
  MultRat \\
  I'm here! \\
  I'm here! \\
  I'm here! \\
  I'm here! \\
  I'm here! \\
  I'm here!
  };

  \node[title]   at (multccircle.north){\uline{$Mult(\C)$}};
  \node[content] at (multccircle.north west){\ \\ \ \\ \ \\
  MultC \\
  I love you all!
  };

  \node[title]   at (multcbox.north){\uline{$Mult(\C)$}};
  \node[content] at (multcbox.north west){\ \\ \ \\ \ \\ 
  MultC \\
  I am over here!
  };

  \node [below=8.6cm, font=\huge] at (atcenter){$\cong$};

  \draw [->] (lfunc)                     -- node [right,midway] {Some text 1} (multratcircle);
  \draw [->] (varieties)                 -- node [above,midway,sloped] {Some text 2} (multratcircle);
  \draw [->] (varieties)                 -- node [right,midway] {Some text 3} (multratbox);
  \draw [->] (multratcircle)             -- node [above,midway] {Some text 4} (multratbox);
  \draw [right hook->] (multratcircle)   -- node [right,midway] {Some text 5} (multccircle);
  \draw [right hook->] (multratbox)      -- node [right,midway] {Some text 6} (multcbox);
  \draw [->,bend right=10] (multcbox)    edge node [above,midway] {Some text 7} (multccircle);
  \draw [->,bend right=10] (multccircle) edge node [below,midway] {Some text 8} (multcbox);

\end{tikzpicture}
}