### 2018-02-16 article.tex

\documentclass[a4paper]{article}

\input{preamble}

\title{New perspectives on multiplicative functions: Lambda-rings and Tannakian symbols}

\author{Ane Espeseth \ \ \ \ \ Andreas Holmstrom \ \ \ \ \ Torstein Vik}

\date{\today}

% Use to only compile one part of the document - don't commit change here
%\includeonly{sections/01-introduction}

\begin{document}

% title page, abstract, table of contents
\include{frontmatter}

% chapters
\include{sections/01-introduction}
\include{sections/02-multiplicativefunctions}
\include{sections/03-adamsalgebras}
\include{sections/04-adamsalgebrasonmultfunc}
\include{sections/05-rationalmultfunc}
\include{sections/06-belltopology}
\include{sections/07-resultsonadamsalgebras}
\include{sections/08-identitiesapproach}
\include{sections/09-symmetricoperations}
\include{sections/10-algebraicgeometry}
\include{sections/11-dirichetletseries}
\include{sections/12-functorial}
\include{sections/13-lambdaringsclassicalfunctions}
\include{sections/14-classesmultfunc}
\include{sections/15-moreexamples}
\include{sections/16-WIP}

% references
\include{references}

\end{document}